Pool Model Paper

### 2023-12-06 pool_model.tex
Corrections to the toxin production section

to be solved with the initial conditions $L(0)=n_0$, $G(0)=0$, and $D(0)=0$, where $n_0$ is the initial bacterial population. 

@@

        L(t) = n_0 e^{-(\mu+\lambda)t}\\
        D(t) = n_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}

@@

It is interesting to see what type of equations result for the total bacterial population $n=L+G+D$:

@@

        \dot{G} = \lambda L + \alpha G\left(1-\frac{n}{N_t}\right)\\
        \dot{n} = - \beta D + \alpha G\left(1-\frac{n}{N_t}\right)

@@

with the initial conditions $G(0)=0$ and $n(0)=n_0$. 

@@

The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $n_0=2G(t_L)$. 

@@

Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $T$. 
It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
In order to take this into account we add an equation for the waste production:

@@

    \dot{L} &=& - \lambda_f R L + \mathcal{B}(R)G\\
    \dot{G} &=& \lambda_f R L +\alpha R G-\mathcal{B}(R)G\\
    \dot{R} &=&-\frac{\alpha}{N_t} R G\\

@@

The function $\mathcal{F}$ denotes the model for the antibiotic production rate. 
We shall discuss two possible models:
    \mathcal{F}_1(R,G)&=&\kappa G\\
    \mathcal{F}_2(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}
We would like to test the following model:

@@

As it can be seen in Figure~\ref{fig:2pool_2sp_toxin}, model $\mathcal{F}_1$ results in $I_1(t)=\kappa\int_{t_0}^tG(t')dt'$. 
As long as $G>0$ toxin is produced, no matter whether the amount of bacteria is still growing. 
Model $\mathcal{F}_2$ results in waste accumulation in steady state $I_2^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
To see this we note $\dot{T}_2=\kappa(\dot{L}+\dot{G})$, which yields: $I_2(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_2(0)=G(0)=0$, $L(0)=n_0$. 
This model assumes that inhibitor is only produced when the bacteria grow and divide.\\

@@

@@

In the latter case one can simplify the equations, given $n_A$ and $n_B$ are non-zero:

>>>>

to be solved with the initial conditions $L(0)=N_0$, $G(0)=0$, and $D(0)=0$, where $N_0$ is the initial bacterial population. 

@@

        L(t) = N_0 e^{-(\mu+\lambda)t}\\
        D(t) = N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}

@@

It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:

@@

        \dot{G} = \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
        \dot{N} = - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)

@@

with the initial conditions $G(0)=0$ and $N(0)=N_0$. 

@@

The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$. 

@@

To introduce active competitive mechanisms, we extend our model with one of the species producing the toxic compounds $T$ during its growth:

@@

@@

The function $\mathcal{F}$ denotes the model for the toxin production rate.
It can be, for example, the accumulation of the bacteria waste which is responsible for the spoilage of food products and affecting another species in negative way.
Consider the system for two species competing for resource $R$ with the production of the toxin $T$ by species $B$.
Assume that the toxin causes the death of the species $A$ with rate $\mu$.
Then the full \ac{ode} system is

@@

Where we consider two possible models of the toxin production:
    \mathcal{F}_1(R,G)&=&\kappa G_B\\
    \mathcal{F}_2(R,G)&=&\kappa\alpha_B R G_B.
Here $\kappa$ is the toxin production rate.
The difference between the two toxin production models for resulting solution of the \acp{ode} is presented in Figure~\ref{fig:2pool_2sp_toxin}.

@@

Model $\mathcal{F}_1$ results in toxin concentration $T_1(t)=\kappa\int_{t_0}^tG(t')dt'$.
That means that as long as there are bacteria in the growth phase $G_B>0$ toxin is produced, no matter whether the total bacterial count $N_B$ is still growing. 
On the other hand, model $\mathcal{F}_2$ results in toxin accumulation in steady state $T_2^s=\lim_{t\to\infty}T_2(t)=\kappa(N_t-N_0)$. 
To see this we note $\dot{T}_2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$. 
This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.\\

@@

In the latter case one can simplify the equations, given $N_A$ and $N_B$ are non-zero:
